---
title: "SingleDataset"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SingleDataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# This vignette will provide guidance for analysis of a single dataset.

Let's assume you have a dataset at the ready. I will try to provide a small example dataset for reproducible tutorials in the future.

first load up the package. 

```{r setup}
#install.packages("devtools") #if devtools not already installed
#devtools::install_github('apf2139/tamlabscpipeline')
library(tamlabscpipeline)
```


Let us assume also that you have access to an output of cellranger in the form of a dir or an H5 file. 

The seuratpipeline() function will perform QC and initial clustering of your dataset. This will involve filtering out high mitochondrial content cells, low library size cells, and doublets; then normalizing using [SingleCellTransform (SCT)](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-019-1874-1) and clustering using the Louvain method as implemented in [Seurat](https://satijalab.org/seurat/).

Here is a quick start for launching the analysis pipeline using this function. I have used the object filepath to store a string containing the path to an example file.

The command prints out a bunch of plots for QC purposes, so best to wrap it in a pdf command block.


```{r seuratpipeline}
#pdf('testrun.pdf')

##sobj <- seuratpipeline(data = filepath, #path to dataset
#               format = h5 #format; dir or H5
               
#               )

```
